Constraints on the mass and occupation fraction of dark matter halos hosting Lyman-alpha emitters at $z$=3.1

### 2013-06-26 paper/ms.tex
abstract + conclusions

\title[Halo mass and occupation fraction for LAEs at $z=3.1$]{Using cosmic
  variance to constrain the dark matter halo mass and occupation
  fraction of Lyman-alpha emitting galaxies at $z$=3.1}

\author[~J.~E. Forero-Romero and J. Mejia]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
We use cosmological N-body dark matter only simulations to constrain the
characteristic mass of dark matter halos hosting Lyman-Alpha Emitting
(LAE) galaxies at a redshift of $z=3.1$. The method is based on
matching the statistics for the number density between mock and
observed fields. The mock fields are constructed using a simple model
where a dark matter halo can only host one LAE with a probability
$f_{\rm occ}$ if its mass is found withing a certain range mass range
delimited by two threshold values, $M_{\rm min}$ and $M_{\rm max}$. 
We also maximize the number of mocks surveys consistent with
Our most important results is that, unlike previous studies, a pure
clustering analysis cannot fully constraint the halo occupation
fraction for LAEs.

Most of the models have a narrow mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}< 1.0$ dex and a low
occupation fraction $f_{\rm occ}\leq 0.3$.

observations and impose consistency with the angular correlation
function. Under these conditions we find that LAEs are preferentially
hosted by halos in very narrow mass ranges less than 0.5 dex in width,
with the minimum mass $10\leq \log_{10}M_{\rm min}\leq 10.9$ and the
occupation fraction $f_{\rm   occ}\leq 0.2$.  Our finding suggest that
the most massive dark matter halos at that epoch do not host the
brightest LAEs. This also gives support to observational evidence that
says that only a small fraction of star forming galaxies can be
actually detected as LAEs.
\end{abstract}

@@

\subsection{Three Regions in Parameter Space}

@@

Without any additional information our method allows us to infer that
most of the succesful models are found in the second and third region of
parameter space where. This result was already expected from halo
abundance calculations shown in Figure \ref{fig:halos}. In the next
section we reduce the size of this region by using more stringent
constraints to define the agreement with observations and including
additional observational information onpossible values for the
occupation fraction at $z\sim 3$.

@@

In this next subsections we consider two additional constraints on the models
we accept as succesul. First we take models with the highest possible
number of mock surveys consistent with observations. Second by
comparing the angular correlation function against observational
results by \cite{Hayashino2004}.

@@

All of the prefered models have a halo mass range with $M_{\rm
  min}<10^{11.4}$\hMsun. This suggests that a detailed and careful
study of the spectral properties of LAEs coupled to the kind of
analysis performed in this paper can be a guide in the study of the
properties of low mass dark matter halos at $z=3.1$, extending the
capabilities of abundance matching methods.

@@

In this \documentname we constrain the preferred dark matter mass
and occupation fraction for halos hosting Lyman Alpha Emitters at
redshift $z=3.1$. We use a method that matches the cosmic variance in the surface
density number of LAEs between mock and real observations and provides
consistent results with observational estimates for the angular
correlation function. The mock catalogs are based on a simplified
model with three basic parameters: the halo mass range where LAEs can
be found, $M_{\rm   min}<M_{\rm h}<M_{\rm   max}$, and the fraction of
the halos in this range that are actully occupied, $f_{\rm occ}$. After
a thorough exploration of the parameter space we find that wide
variety of models are consistent with the observational constraints,
where only a minority has been considered so far in the
literature. Out of $9000$ initial combinations for the model
parameters we end up with $49$ sucessful arrangement of parameters.

These parameter sets can be split into three families depending the mass range
$\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm min}$ and the
occupation fraction $f_{\rm occ}$. The first family is narrow both in
$\Delta M$ and $f_{\rm occ}$, a second familiy is only narrow in
$\Delta M$ and the third is narrow in $f_{\rm occ}$ and broad in
$\Delta M$.

All the halo mass ranges in the best models fall into the broad
results based in observations, which have been able to describe models
in the third family, but have so far missed all the possible
models. In particular, previous analysis did not consider the first
two families: a narrow mass range and any value for the escape fraction.

In the case of narrow mass models there are various cases with $\Delta
M=0.1$dex. Taken at face value, a possible confirmation of these kind of models
can be considered a challenge to explain relatively high star
formation ratesm

\section*{Acknowledgments}

>>>>

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Constraints on the mass and occupation
  fraction of dark matter halos hosting Lyman-alpha emitters
  at $z$=3.1}    
\author[~J.~E. Forero-Romero and J. Mejia]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$, by comparing the  number density and the angular
correlation function between mock and observed fields. The mock fields
are constructed in a large cosmological N-body simulation using a
model where a dark matter halo can only host one LAE with a
probability $f_{\rm occ}$ if its mass is found within the mass range
delimited by two threshold values  $M_{\rm min}$ and $M_{\rm max}$. We
conclude that the number density and spatial clustering information
available are not sufficient to derive a unique and narrow set of
values for the parameter in the model. We find three different
families of models based on their mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}< 1.0$ dex and occupation
fraction $f_{\rm occ}$.  The dominant familiy is composed by models with a narrow
mass range $\Delta M<1.0$ dex, a low occupation fraction $f_{\rm occ}\leq 0.3$
and a maximum mass $M_{\rm max}< 10^{12}$\hMsun.  This finding
suggests that the most massive dark matter halos at that epoch do not
host the brightest LAEs. This also gives support to observational evidence
concluding that only a small fraction of star forming galaxies can be
actually detected as LAEs. We provide tables and mock catalogs for the
succesful models.

\end{abstract}

@@

\subsection{Three halo mass ranges}

@@

Without any additional information our method allows us to infer that
most of the succesful models are found in the second and third region of
parameter space where. This result was already expected from halo
abundance calculations shown in Figure \ref{fig:halos}. In the next
section we reduce the size of this region by using more stringent
constraints to define the agreement with observations and the
inclusion of the angular correlation function information by
\cite{Hayashino2004}.

@@

@@

All the prefered models have a halo mass range with $M_{\rm
  min}<10^{11.4}$\hMsun. This suggests that a detailed and careful
study of the spectral and photometric properties of LAEs coupled to the kind of
analysis performed in this paper can be a guide in the study of the
properties of low mass dark matter halos at $z=3.1$, extending the
capabilities of abundance matching methods.

@@

In this \documentname we constrain the preferred dark matter mass
and occupation fraction for halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We use a method that
matches the cosmic variance of the surface 
density number of LAEs between mock and real observations and is also
consistent with observational estimates for the angular
correlation function. The mock catalogs are constructed using a 
model with three basic parameters: the halo mass range where LAEs can
be found, $M_{\rm   min}<M_{\rm h}<M_{\rm   max}$, and the fraction of
the halos in this range that are actully occupied, $f_{\rm occ}$. After
a thorough exploration of the parameter space we find that wide
variety of models are consistent with the observational constraints,
where only a minority has been considered so far in the
literature. Out of $9000$ initial combinations for the model
parameters we end up with $49$ sucessful arrangement of parameters.

The most important conclusion of this work is that available
observational information for the clustering properties of LAEs at
$z=3.1$ does not provide enough constraints to uniquely narrow the range of masses
and occupation fraction of dark matter halos hosting those galaxies. In our
model, the hosting halos can be split into three families depending the mass range
$\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm min}$ and the
occupation fraction $f_{\rm occ}$. The first family is narrow both in
$\Delta M$ and $f_{\rm occ}$, a second familiy is only narrow in
$\Delta M$ and the third is narrow in $f_{\rm occ}$ and broad in
$\Delta M$.

All the halo mass ranges in the best models fall into the broad
range obtained in previous analysis \citep{Gawiser2007}. The
improvement of our method is being able to describe in high detail
all the possible models consistent with observations, including the
discovery of models previously overlooked like those with a high
occupation fraction $f_{\rm occ}>0.3$.

A central result in this exploration of paremter espace is the
existence of a large number of models with a very
narrow mass range $\Delta M< 0.1$ dex that are consistent in every
aspect with the spatial distribution of observed LAEs. In order to
have these characteristics it is required that halos around the mass
$M_{\rm max}$ for each model become ineffecient in hosting detectable
galaxies. One way to achieve this is having LAEs with a decreasing \ly
escape fraction with increasing mass.

All the models also present a range of masses where the halos are
below the imass scale of $10^{11.4}$\hMsun infered as a lower
threshold for the LBGs used in halo abundace matching
studies. Detailed study of high-z LAES will have a possitive impact on
our undertading of the connection of low mass halos to star forming
galaxies.

We forsee that the new observations with new instruments (such as MUSE,
Hyper SuprimeCam and HETDEX) covering larger fields and a wider range
of luminosities will be key in imposing tighter constraints on the
properties of dark matter halos hosting LAEs.

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advance Studies where the first full draft of this paper
was completed.
